---
title: "Bike"
author: "Dello Russo, Deviardi, Gorbani"
date: "2025-02-11"
output:
  
  html_document: 
    df_print: paged
    toc: true
  pdf_document: default
  html_notebook: 
    df_print: paged
  word_document: default
editor_options: 
  chunk_output_type: console
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(rgl)
knitr::opts_chunk$set(
  echo = TRUE,        # Mostra il codice
  results = "hide"   # Nasconde i grafici
)
knitr::knit_hooks$set(webgl = hook_webgl)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
ggplot2::theme_set(ggplot2::theme_bw())
```


```{css, echo=FALSE}
.extracode {
background-color: lightblue;
}
```

# Libraries

```{r}
library(dplyr)
library(tidyr)
library(kableExtra)
library(corrplot)
library(fdatest)
library(data.table)
library(roahd)
library(mgcv)
library(ggplot2)
library(knitr)
library(broom)
library(tidyr)
library(progress)
library(pbapply)
library(dbscan)
library(gridExtra)
library(fda)
library(roahd)
library(conformalInference.fd)
library(quantreg)
library(splines)
library(rgl)

```

# Dataset

```{r}
day <- read.csv("Dataset/day.csv")
hour <- read.csv("Dataset/hour.csv")

# checking the number of observations in each day
blocchi <- rle(hour$dteday)
blocchi$values   
blocchi$lengths

# creating the variables of cnt_hour_registered in the day dataset
hour_wide_register <- hour %>%
  pivot_wider(
    names_from = hr, 
    values_from = registered,
    names_prefix = "cnt_hour_"
  )
hour_wide_sommato_reg <- hour_wide_register %>%
  group_by(dteday) %>%
  summarise(across(starts_with("cnt_hour_"), sum, na.rm = TRUE))
day2<- day %>%
  left_join(hour_wide_sommato_reg, by = "dteday")

hour_wide_casual <- hour %>%
  pivot_wider(
    names_from = hr, 
    values_from = casual,
    names_prefix = "cnt_hour_"
  )
hour_wide_sommato_cas <- hour_wide_casual %>%
  group_by(dteday) %>%
  summarise(across(starts_with("cnt_hour_"), sum, na.rm = TRUE))
day1 <- day %>%
  left_join(hour_wide_sommato_cas, by = "dteday")

#Trasforma `dati_orari` per avere una colonna per ogni ora
hour_wide <- hour %>%
  pivot_wider(
    names_from = hr, 
    values_from = cnt,
    names_prefix = "cnt_hour_"
  )
hour_wide_sommato <- hour_wide %>%
  group_by(dteday) %>%
  summarise(across(starts_with("cnt_hour"), sum, na.rm = TRUE))
day <- day %>%
  left_join(hour_wide_sommato, by = "dteday")

kable(head(day[,1:16]), caption = "Prime righe del dataset day") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = F)

```

# Exploratory Data Analysis

```{r}
# Outliers--------
ind_outlier<-which(blocchi$lengths==1)
day[ind_outlier,] # 2012-10-29: Hurricane Sandy
day<-day[-ind_outlier,]
day1<-day1[-ind_outlier,]
day2<-day2[-ind_outlier,]

# Graph---------
# Hourly rentals
matplot(t(day[, 17:40]), 
        type = 'l', 
        col = ifelse(day$workingday == 1, 'blue', 
                     ifelse(day$holiday == 1, 'red', 'green')),xlab="hour", ylab="users")
legend("topleft", legend=c("working day","weekend", "holiday"), col=c("blue","green","red"), lty=1, lwd=2, cex=0.75)
# Daily rentals
plot(day$instant,day$cnt,xlab="day",ylab="user")
# Rentals by weather situation
ggplot(day, aes(x = as.factor(weathersit), y = cnt)) +
  geom_boxplot() +
  labs(title = "Rentals by Weather Situation", x = "Weather Situation", y = "Total Rentals")
# Corrplot
meteo<-day[,10:16]
cor_meteo<-cor(meteo)
corrplot(cor_meteo, method = "number")
# Register and casual
## hour
plot(seq(0,23),colMeans(day2[,17:40]),type="l",xlab="hour",ylab="cnt",col='red')
lines(seq(0,23),colMeans(day1[,17:40]),col='blue')
legend("topleft", legend=c("registered","casual"), col=c("red","blue"), lty=1, lwd=2, cex=0.75)

## day
plot(day$instant,day$registered,xlab="day",ylab="rentals",col='red')
points(day$instant,day$casual,col='blue')
legend("topleft", legend=c("registered","casual"), col=c("red","blue"), lty=1, lwd=2, cex=0.75)

```

# Registered vs casual

```{r}
t1 = day1[,17:40]
t2 = day2[,17:40]

t1.mean = colMeans(t1)
t2.mean = colMeans(t2)

matplot(seq(0,23),
        t(rbind(t1.mean,t2.mean)), 
        type='l', col=c(1,2), lty=1,
        main="Sample multivariate means")

n1 = dim(t1)[1]
n2 = dim(t2)[1]
n  = n1 + n2

T20 = as.numeric(t(t1.mean-t2.mean) %*% (t1.mean-t2.mean))  # matrix (vector) product
#T20 <- norm(t1.mean-t2.mean) same as before
T20
B<-1000
T2 = numeric(B)
set.seed(19)
for(perm in 1:B){
  # Random permutation of indexes
  # When we apply permutations in a multivariate case, we keep the units together
  # i.e., we only permute the rows of the data matrix
  t_pooled = rbind(t1,t2)
  permutation = sample(n)
  t_perm = t_pooled[permutation,]
  t1_perm = t_perm[1:n1,]
  t2_perm = t_perm[(n1+1):n,]
  
  # Evaluation of the test statistic on permuted data
  t1.mean_perm = colMeans(t1_perm)
  t2.mean_perm = colMeans(t2_perm)
  T2[perm]  = t(t1.mean_perm-t2.mean_perm) %*% (t1.mean_perm-t2.mean_perm) 
}
hist(T2,xlim=range(c(T2,T20)))
abline(v=T20,col=3,lwd=4)
p_val = sum(T2>=T20)/B
p_val#0
```



# Registered


## Hour usage

### Permutational test working day - weekend distribution

```{r}
a<-subset(day2,day2$workingday==1)
matplot(t(a[,17:40]), type='l')
b<-subset(day2,day2$workingday==0)
matplot(t(b[,17:40]), type='l')
x<- a$cnt
y<-b$cnt
perm_t_test=function(x,y,iter=1e3){
  T0=abs(mean(x)-mean(y))  # define the test statistic
  T_stat=numeric(iter) # a vector to store the values of each iteration
  x_pooled=c(x,y) # pooled sample
  n=length(x_pooled)
  n1=length(x)
  print(sprintf("Using %s iterations to estimate the permutational distribution. There are actually %s possible permutations", iter, factorial(n)))
  for(perm in 1:iter){ # loop for conditional MC
    # permutation:
    permutation <- sample(1:n)
    x_perm <- x_pooled[permutation]
    x1_perm <- x_perm[1:n1]
    x2_perm <- x_perm[(n1+1):n]
    # test statistic:
    T_stat[perm] <- abs(mean(x1_perm) - mean(x2_perm))
  }
  # p-value
  p_val <- sum(T_stat>=T0)/iter
  return(p_val)
}
p.value <- perm_t_test(x,y,iter=1e3) #0.074
```

### Interval-wise Test

```{r}
fd<-fData(seq(0,23),(day2[,17:40]))
day_hour <- as.matrix(day2[,17:40])
groups <- day$workingday
ITP.result <- ITPaovbspline(day_hour ~ groups, B=1000,nknots=20,order=3,method="responses") #ITPlmbspline per la regressione
summary(ITP.result)
plot(ITP.result,plot.adjpval = TRUE, xrange=c(0,23), )
```

### Conformal prediction

```{r}
day2011= subset(day2, day$yr==0)
day2012= subset(day2, day$yr==1)

a<-subset(day2,day2$workingday==1)
a2011 = subset(day2011,day2011$workingday==1)
a2012 = subset(day2012,day2012$workingday==1)

b<-subset(day2,day2$workingday==0)
b2011 = subset(day2011,day2011$workingday==0 & day2011$holiday==0)
b2012 = subset(day2012,day2012$workingday==0 & day2012$holiday==0)

par(mfrow=c(1,2))
matplot(t(a2012[,17:40]), type='l', col='darkorange', ylim = c(0,1100), main = "2012 working day")
matplot(t(b2012[,17:40]), type='l', col = 'forestgreen', ylim = c(0,1100), main= "2012 weekend")

# 2012 workingday 5-20
dataset <- a2012
x.grid <- 5:20
f_data <- fData(x.grid, dataset[,22:37])
rho <- 0.5 # train and calibration split ratio

alpha = 0.1

l <- dim(dataset)[1] - ceiling(dim(dataset)[1] * rho)
alpha_rag <- c(seq(1/(l+1), l/(l+1), 1/(l+1)))

fun=mean_lists()
x0=list(as.list(x.grid))

band <- conformal.fun.split(NULL, NULL, f_data, NULL, x0, fun$train.fun, fun$predict.fun,
                            alpha = alpha,
                            split=NULL, seed=2024, randomized=FALSE, seed.rand=FALSE,
                            verbose=FALSE, rho=rho, s.type="st-dev")

upper_b_a <- band$up[[1]][[1]]
lower_b_a <- band$lo[[1]][[1]]
point_pred_a <- band$pred[[1]][[1]]

# 2012 weekend 5-20
dataset <- b2012
x.grid <- 5:20
f_data <- fData(x.grid, dataset[,22:37])
rho <- 0.5 # train and calibration split ratio

alpha = 0.1

l <- dim(dataset)[1] - ceiling(dim(dataset)[1] * rho)
alpha_rag <- c(seq(1/(l+1), l/(l+1), 1/(l+1)))

fun=mean_lists()
x0=list(as.list(x.grid))

band <- conformal.fun.split(NULL, NULL, f_data, NULL, x0, fun$train.fun, fun$predict.fun,
                            alpha = alpha,
                            split=NULL, seed=2024, randomized=FALSE, seed.rand=FALSE,
                            verbose=FALSE, rho=rho, s.type="st-dev")

upper_b_b <- band$up[[1]][[1]]
lower_b_b <- band$lo[[1]][[1]]
point_pred_b <- band$pred[[1]][[1]]


par(mfrow=c(1,2))
plot(x.grid, upper_b_a, type='l', col="black", lty=2, xlab="hour", ylab="cnt", lwd=1.5,
     ylim=c(min(lower_b_b),max(upper_b_a)),
     main="Functional CP band working day")
lines(x.grid, point_pred_a, type='l', col="darkorange", lwd=1.5)
lines(x.grid, lower_b_a, type='l', col="black", lty=2, lwd=1.5)

plot(x.grid, upper_b_b, type='l', col="black", lty=2, xlab="hour", ylab="cnt", lwd=1.5,
     ylim=c(min(lower_b_b),max(upper_b_a)),
     main="Functional CP band weekend")
lines(x.grid, point_pred_b, type='l', col="forestgreen", lwd=1.5)
lines(x.grid, lower_b_b, type='l', col="black", lty=2, lwd=1.5)

par(mfrow=c(1,1))

```

## Weather 

### Permutational test: regression

```{r}
n<- dim(day2)[1]
result<-lm(registered ~ hum +temp + windspeed+instant,data=day2)
summary(result)
shapiro.test(result$residuals)$p
qqnorm(result$residuals)
qqline(result$residuals)
#Let’s start with a global test
B<-1000
T0_glob <- summary(result)$adj.r.squared
T_H0glob <- numeric(B)
for(perm in 1:B){
  permutation <- sample(n)
  
  Y.perm.glob <- with(day2,registered[permutation])
  T_H0glob[perm] <- summary(lm(Y.perm.glob ~ hum +temp +instant+ windspeed,data=day2))$adj.r.squared
}
sum(T_H0glob>=T0_glob)/B#0

T0_x1 <- abs(summary(result)$adj.r.squared)
T0_x2 <- abs(summary(result)$adj.r.squared)
T0_x3 <- abs(summary(result)$adj.r.squared)
T0_x4<- abs(summary(result)$adj.r.squared)
regr.H01 <- lm(registered~temp+windspeed+instant,data=day2)
residuals.H01 <- regr.H01$residuals

regr.H02 <- lm(registered~hum+windspeed+instant,data=day2)
residuals.H02 <- regr.H02$residuals

regr.H03 <- lm(registered~hum+temp+instant,data=day2)
residuals.H03 <- regr.H03$residuals

regr.H04 <- lm(registered~hum+temp+windspeed,data=day2)
residuals.H04 <- regr.H04$residuals

T_H01 <- T_H02 <- T_H03 <- T_H04<- numeric(B)
set.seed(202454)
for(perm in 1:B){
  permutation <- sample(n)
  
  residuals.H01.perm <- residuals.H01[permutation]
  Y.perm.H01 <- regr.H01$fitted + residuals.H01.perm
  T_H01[perm] <- abs(summary(lm(Y.perm.H01 ~ hum+temp+windspeed+instant,data=day2 ))$adj.r.squared)
  
  residuals.H02.perm <- residuals.H02[permutation]
  Y.perm.H02 <- regr.H02$fitted + residuals.H02.perm
  T_H02[perm] <- abs(summary(lm(Y.perm.H02 ~ hum+temp+windspeed+instant,,data=day2))$adj.r.squared)
  
  residuals.H03.perm <- residuals.H03[permutation]
  Y.perm.H03 <- regr.H03$fitted + residuals.H03.perm
  T_H03[perm] <- abs(summary(lm(Y.perm.H03 ~ hum+temp+windspeed+instant,,data=day2))$adj.r.squared)
  
  residuals.H04.perm <- residuals.H04[permutation]
  Y.perm.H04 <- regr.H04$fitted + residuals.H04.perm
  T_H04[perm] <- abs(summary(lm(Y.perm.H04 ~ hum+temp+windspeed+instant,,data=day2))$adj.r.squared)
}

sum(T_H01>=T0_x1)/B#0.015
sum(T_H02>=T0_x2)/B#0
sum(T_H03>=T0_x3)/B#0.183
sum(T_H04>=T0_x4)/B#0
```

### trend
```{r}
nbasis <- 10 
bspline_basis <- create.bspline.basis(rangeval = c(1,730), nbasis = nbasis)

lambda <- 1  
fdParobj <- fdPar(bspline_basis, lambda = lambda)
smoothed_fd <- smooth.basis(1:730, day2$registered, fdParobj)$fd
func <- eval.fd(1:730, smoothed_fd)

plot(1:730, day$registered, pch = 16, col = "gray", main = "Smoothing con Basi di Fourier",
     xlab = "Tempo", ylab = "Valore")
lines(1:730, func, col = "blue", lwd = 2)
legend("topright", legend = c("Dati", "Smoothing"),
       col = c("gray", "blue"), lty = c(NA, 1, 2), pch = c(16, NA, NA))
stl_result <- stl(ts(func[,1], frequency = 364), s.window = "periodic")  
plot(stl_result)

```

### Semiparametric regression

```{r}
model_gam_inter=gam(registered ~ instant + s(hum,bs='cr')+s(temp,bs='cr')+s(I(hum*temp),bs='cr'),data = day2)
summary(model_gam_inter)

model_gam_reduced=gam(registered ~ instant + s(hum,bs='cr')+s(temp,bs='cr'),data = day)
summary(model_gam_reduced)

model_gam_tp = gam(registered ~ instant + s(hum, temp, bs="tp", m=2), # m for order
                   data = day)

temp.grid=(seq(range(day$temp)[1],range(day$temp)[2],by=0.01))
hum.grid=(seq(range(day$hum)[1],range(day$hum)[2],by=0.01))
ist.grid=(seq(range(day$instant)[1],range(day$instant)[2],by=1))
grid=expand.grid(instant = 365, hum = hum.grid, temp = temp.grid)
grid=as.data.frame(grid)
pred_tp = predict(model_gam_reduced,
                  newdata = grid)

open3d()
persp3d(hum.grid,temp.grid, pred_tp, col = 'cyan')
points3d(day$hum, day$temp, day$cnt, col = 'black', size = 5)

hum.grid <- seq(min(day$hum), max(day$hum), length = 30)  # 30 punti tra min e max di hum
temp.grid <- seq(min(day$temp), max(day$temp), length = 30)  # 30 punti tra min e max di temp

# Creare una griglia combinando tutti i valori di hum e temp
grid <- expand.grid(hum = hum.grid, temp = temp.grid, instant = 1)
pred_tp <- predict(model_gam_reduced, newdata = grid)
pred_tp_matrix <- matrix(pred_tp, nrow = length(hum.grid), ncol = length(temp.grid))
open3d()
persp3d(hum.grid, temp.grid, pred_tp_matrix, col = 'grey30', alpha = 0.7, 
        xlab = "Humidity", ylab = "Temperature", zlab = "Bike Rentals")

# Aggiungere i punti reali in rosso
points3d(day$hum, day$temp, day$cnt, col = 'red', size = 5)
predict(model_gam_reduced, newdata = data.frame(hum = 0.51, temp = 11/41,instant=739))
predict(model_gam_inter, newdata = data.frame(hum = 0.51, temp = 11/41,instant=739))

anova(model_gam_tp ,model_gam_inter, test = "F")
shapiro.test(model_gam_tp$residuals)


```


# Casual
### outliers

```{r}
fCasual<-fData(seq(0,23),(day1[,17:40]))
fbplot <- invisible(fbplot(fCasual, main="Magnitude outliers"))
fbplot$ID_outliers
day1[fbplot$ID_outliers,]
day1<-day1[-185,]
day1<-day1[-551,]
hist(day1$casual)
holiday_index<-which(day1$holiday==1)
day1[holiday_index,]
day1<-day1[-holiday_index,]
```


### Permutational test: regression

```{r}
n<- dim(day1)[1]
result<-lm(casual ~ hum +temp + windspeed+instant,data=day1)
summary(result)
shapiro.test(result$residuals)$p
qqnorm(result$residuals)
qqline(result$residuals)
#Let’s start with a global test
B<-1000
T0_glob <- summary(result)$adj.r.squared
T_H0glob <- numeric(B)
for(perm in 1:B){
  permutation <- sample(n)
  
  Y.perm.glob <- with(day1,casual[permutation])
  T_H0glob[perm] <- summary(lm(Y.perm.glob ~ hum +temp + windspeed+instant,data=day1))$adj.r.squared
}
sum(T_H0glob>=T0_glob)/B#0

T0_x1 <- abs(summary(result)$adj.r.squared)
T0_x2 <- abs(summary(result)$adj.r.squared)
T0_x3 <- abs(summary(result)$adj.r.squared)
T0_x4 <- abs(summary(result)$adj.r.squared)

regr.H01 <- lm(casual~temp+windspeed+instant,data=day1)
residuals.H01 <- regr.H01$residuals

regr.H02 <- lm(casual~hum+windspeed+instant,data=day1)
residuals.H02 <- regr.H02$residuals

regr.H03 <- lm(casual~hum+temp+instant,data=day1)
residuals.H03 <- regr.H03$residuals

regr.H04 <- lm(casual~hum+temp+weathersit,data=day1)
residuals.H04 <- regr.H04$residuals

T_H01 <- T_H02 <- T_H03 <-T_H04<- numeric(B)

for(perm in 1:B){
  permutation <- sample(n)
  
  residuals.H01.perm <- residuals.H01[permutation]
  Y.perm.H01 <- regr.H01$fitted + residuals.H01.perm
  T_H01[perm] <- abs(summary(lm(Y.perm.H01 ~ hum+temp+windspeed+instant,data=day1))$adj.r.squared)
  
  residuals.H02.perm <- residuals.H02[permutation]
  Y.perm.H02 <- regr.H02$fitted + residuals.H02.perm
  T_H02[perm] <- abs(summary(lm(Y.perm.H02 ~ hum+temp+windspeed+instant,data=day1))$adj.r.squared)
  
  residuals.H03.perm <- residuals.H03[permutation]
  Y.perm.H03 <- regr.H03$fitted + residuals.H03.perm
  T_H03[perm] <- abs(summary(lm(Y.perm.H03 ~ hum+temp+windspeed+instant,data=day1))$adj.r.squared)
  
  residuals.H04.perm <- residuals.H04[permutation]
  Y.perm.H04 <- regr.H04$fitted + residuals.H04.perm
  T_H04[perm] <- abs(summary(lm(Y.perm.H04 ~ hum+temp+windspeed+instant,data=day1))$adj.r.squared)
}

sum(T_H01>=T0_x1)/B#0.14
sum(T_H02>=T0_x2)/B#0
sum(T_H03>=T0_x3)/B#0.389
sum(T_H04>=T0_x4)/B#0.035
```


### Quantile regression

```{r}
fit_rq<-rq(casual~temp,data=day1,tau=c( .5))
day1$res_rq<-fit_rq$residuals
a<-subset(day1,day1$workingday==0 | day1$weekday==5) # dataset friday and weekend
b<-subset(day1,day1$workingday==1 & day1$weekday!=5) # dataset Mon-Thurs
x<- a$res_rq
y<-b$res_rq
perm_t_test=function(x,y,iter=1e3){
  T0=abs(mean(x)-mean(y))  # define the test statistic
  T_stat=numeric(iter) # a vector to store the values of each iteration
  x_pooled=c(x,y) # pooled sample
  n=length(x_pooled)
  n1=length(x)
  print(sprintf("Using %s iterations to estimate the permutational distribution. There are actually %s possible permutations", iter, factorial(n)))
  for(perm in 1:iter){ # loop for conditional MC
    # permutation:
    permutation <- sample(1:n)
    x_perm <- x_pooled[permutation]
    x1_perm <- x_perm[1:n1]
    x2_perm <- x_perm[(n1+1):n]
    # test statistic:
    T_stat[perm] <- abs(mean(x1_perm) - mean(x2_perm))
  }
  # p-value
  p_val <- sum(T_stat>=T0)/iter
  return(p_val)
}
p.value <- perm_t_test(x,y,iter=1e3) #0

fit_s<-rq(casual~temp,data=a,tau=c( .5))
a$res_rq<-fit_s$residuals
x<- a$res_rq[a$workingday==1] # Friday
y<-a$res_rq[a$workingday==0] # Weekend
perm_t_test=function(x,y,iter=1e3){
  T0=abs(mean(x)-mean(y))  # define the test statistic
  T_stat=numeric(iter) # a vector to store the values of each iteration
  x_pooled=c(x,y) # pooled sample
  n=length(x_pooled)
  n1=length(x)
  print(sprintf("Using %s iterations to estimate the permutational distribution. There are actually %s possible permutations", iter, factorial(n)))
  for(perm in 1:iter){ # loop for conditional MC
    # permutation:
    permutation <- sample(1:n)
    x_perm <- x_pooled[permutation]
    x1_perm <- x_perm[1:n1]
    x2_perm <- x_perm[(n1+1):n]
    # test statistic:
    T_stat[perm] <- abs(mean(x1_perm) - mean(x2_perm))
  }
  # p-value
  p_val <- sum(T_stat>=T0)/iter
  return(p_val)
}
p.value <- perm_t_test(x,y,iter=1e3) #0

plot(day1$temp,day1$casual,cex=.25,type="n",xlab="Temperature", ylab="Rentals")
points(day1$temp[day1$workingday==0],day1$casual[day1$workingday==0],cex=.5,col='red')
points(day1$temp[day1$weekday==5],day1$casual[day1$weekday==5],cex=.5,col='blue')
points(day1$temp[day1$workingday==1 & day1$weekday!=5],day1$casual[day1$workingday==1 & day1$weekday!=5],cex=.5,col='green')
legend("topleft", legend=c("Weekend","Friday","Mon-Thurs"), col=c("red","blue","green"), pch=1,)

```

#### Quantile regression Weekend

```{r}
day1_weekend<-subset(day1,day1$workingday==0)

fit_rq<-rq(casual~temp,data=day1_weekend,tau= .5)
day1_weekend$res<-fit_rq$residuals
x<- day1_weekend$res[day1_weekend$weekday==6]
y<-day1_weekend$res[day1_weekend$weekday==0]
perm_t_test=function(x,y,iter=1e3){
  T0=abs(mean(x)-mean(y))  # define the test statistic
  T_stat=numeric(iter) # a vector to store the values of each iteration
  x_pooled=c(x,y) # pooled sample
  n=length(x_pooled)
  n1=length(x)
  print(sprintf("Using %s iterations to estimate the permutational distribution. There are actually %s possible permutations", iter, factorial(n)))
  for(perm in 1:iter){ # loop for conditional MC
    # permutation:
    permutation <- sample(1:n)
    x_perm <- x_pooled[permutation]
    x1_perm <- x_perm[1:n1]
    x2_perm <- x_perm[(n1+1):n]
    # test statistic:
    T_stat[perm] <- abs(mean(x1_perm) - mean(x2_perm))
  }
  # p-value
  p_val <- sum(T_stat>=T0)/iter
  return(p_val)
}
p.value <- perm_t_test(x,y,iter=1e3)#0.125

temp<-day1_weekend$temp
plot(temp,day1_weekend$casual,cex=.25,type="n",xlab="Temperature", ylab="Rentals")
col_saturaday=ifelse(day1_weekend$weekday==6,"green","cyan")
points(temp,day1_weekend$casual,cex=.5,col=col_saturaday)
legend("topleft", legend=c("Saturday","Sunday"), col=c("green","cyan"), pch=1,)
abline(rq(casual~temp,tau=.5,data=day1_weekend),col="black") 
abline(lm(casual~temp,data=day1_weekend),lty=2,col="red") 
taus <- c(.05,.1,.25,.75,.90,.95)

for( i in 1:length(taus)){
  abline(rq(casual~temp,tau=taus[i],data=day1_weekend),col="gray")
}

fit1 <- rq(casual~temp,tau=.25,data=day1_weekend)
fit2 <- rq(casual~temp,tau=.50,data=day1_weekend)
fit3 <- rq(casual~temp,tau=.75,data=day1_weekend)
anova(fit1, fit2, fit3)


```

#### Quantile regression Friday

```{r}
friday<-subset(day1,day1$weekday==5)

with(friday,plot(temp,casual,cex=.25,type="n",xlab="Temperature", ylab="Rentals",ylim = c(0,3500)))
with(friday,points(temp,casual,cex=.5))
abline(rq(casual~temp,tau=.5,data=friday),col="black") 
abline(lm(casual~temp,data=friday),lty=2,col="red") 
taus <- c(.05,.1,.25,.75,.90,.95)

for( i in 1:length(taus)){
  abline(rq(casual~temp,tau=taus[i],data=friday),col="gray")
}
fit1 <- rq(casual~temp,tau=.25,data=friday)
fit2 <- rq(casual~temp,tau=.50,data=friday)
fit3 <- rq(casual~temp,tau=.75,data=friday)
anova(fit1, fit2, fit3)

```

#### Quantile regression Mon-Thurs

```{r}
day_work <- b
gio <- ifelse(day_work$weekday==4,1,0)
mer <- ifelse(day_work$weekday==3,1,0)
fit2 <- rq(casual ~ temp, data = day_work, tau = 0.5)
x<- fit2$residuals[mer==1]
y<-fit2$residuals[mer==0]
perm_t_test(x,y,iter=1e3) 
plot(day_work$temp, day_work$casual)
which(day_work$casual>1600)
day_work$t <- 1:416 #eliminati outlier in riga 523 (holiday)
colours <- rainbow(4)
col <- colours[day_work$weekday]
plot(day_work$temp,day_work$casual, col=col, xlab="Temperature", ylab="Rentals",ylim=c(0,3500))
names <- c("Monday","Tuesday","Wednesday","Thursday")
legend("topleft", legend=names, col=col, pch=1)
abline(rq(casual~temp,data=day_work,tau=.5),col="blue") 
abline(lm(casual~temp,data=day_work),lty=2,col="red") #the dreaded ols line
taus <- c(.05,.1,.25,.75,.90,.95)

for( i in 1:length(taus)){
 abline(rq(casual~temp,data=day_work,tau=taus[i]),col="gray")
}
```


